---
title: "CodeBook"
author: "omar gonzáles díaz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---




```{r echo=FALSE, results='hide', message=FALSE, warning=F}

library(dplyr)
library(reshape2)
library(pander)

setwd("D:\\RCoursera\\Getting and cleaning data\\UCI HAR Dataset")

features <- readLines("features.txt")

activities <- readLines("activity_labels.txt")


### TEST DIRECTORI

setwd("D:\\RCoursera\\Getting and cleaning data\\UCI HAR Dataset\\test")



test.subject <- read.table("subject_test.txt")


x.test <- read.table("X_test.txt")
y.test <- readLines("y_test.txt")





test <- cbind(test.subject, y.test, x.test)

colnames(test)[1] <- "id"

colnames(test)[2] <- "activities"

colnames(test)[3:563] <- features



setwd("D:\\RCoursera\\Getting and cleaning data\\UCI HAR Dataset\\train")





train.subject <- read.table("subject_train.txt")


x.train <- read.table("X_train.txt")
y.train <- readLines("y_train.txt")


train <- cbind(train.subject, y.train, x.train)

colnames(train)[1] <- "id"

colnames(train)[2] <- "activities"

colnames(train)[3:563] <- features





### Mergging the 2 data frames:

proyect <- rbind(train, test)

proyect$activities <- as.factor(proyect$activities)

### Get all proyect names of means and std, besides "id" and "activities"

proyect.names <- colnames(proyect)


mean.and.std <- grep("mean|std", proyect.names, value = F)
id.activity <- grep("^id$|^activities$", proyect.names, value = F)

### Keeping only relevant columns

proyect <- proyect[,c(id.activity,mean.and.std)]

### Class of "activities" is factor, nees to be "numeric"

proyect$activities <- as.numeric(as.character(proyect$activities))


### Reeplacing code numbers with their meaning: Activities.


for (i in 1:nrow(proyect)) {
        
        
        if(proyect$activities[i] == 1) {
                proyect$activities[i] <- "WALKING"  
        }
        
        else if (proyect$activities[i] == 2) {
                proyect$activities[i] <- "WALKING_UPSTAIRS"
        }
        
        else if (proyect$activities[i] == 3) {
                proyect$activities[i] <- "WALKING_DOWNSTAIRS"
        }
        
        else if (proyect$activities[i] == 4) {
                proyect$activities[i] <- "SITTING"
        }
        
        else if (proyect$activities[i] == 5) {
                proyect$activities[i] <- "STANDING"
        }
        
        else if (proyect$activities[i] == 6) {
                proyect$activities[i] <- "LAYING"
        }
        
        else {
                proyect$activities[i] <- proyect$activities[i]
        }
}


##### STEP 4: renaming variables to make them easier to work with

new.features <- colnames(proyect)


new.features <- sub("bodybody", "Body", new.features, ignore.case = T)
new.features <- sub("[0-9]{1,3} ", "", new.features, ignore.case = T)
new.features <- gsub("[-]*", "", new.features, ignore.case = T)
new.features <- gsub("[()]*", "", new.features, ignore.case = T)
new.features <- tolower(new.features)


colnames(proyect) <- new.features


### Step 5: Making an independant tidy data set

#We need to take the out what our "measure.vars" are going to be

measurevars <- new.features[3:length(new.features)]

#Transform from wide to long (tidy data)

proyect2 <- melt(proyect,
                  # ID variables - all the variables to keep but not split apart on
                  id.vars=c("id", "activities"),
                  # The source columns
                  measure.vars= measurevars,
                  # Name of the destination column that will identify the original
                  # column that the measurement came from
                  variable.name="features",
                  value.name="value")




proyect2$id <- as.factor(proyect2$id)
proyect2$activities <- as.factor(proyect2$activities)


```




## Project Description

This is the code book for the "Getting and Cleaning data course" from the Coursera Data Science Specialization.

##Study design and data processing

###Collection of the raw data

We have worked with the data provided by the course. If you need to find the original data, it comes from:

[Original data source](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones)


###Notes on the original (raw) data 

The original data contained some misspelling errors. Also they contained not so correct variable names which were changed according to the steps suggested in the course:

- All lowercase when possible.
- Descriptive.
- Not duplicated.
- No underscores, dots or white spaces.

##Creating the tidy datafile

###Guide to create the tidy data file

Description on how to create the tidy data file:

1. Load the following libraries: dplyr and reshape2.
2. Load the data provided: test and training data.
3. Renaming the column names.
4. Renaming the column activities with the real activities.
5. Renaming variables to make them easier to work with.
6. Creating the tidy data set.


###Cleaning of the data


##Description of the variables in the tiny_data.txt file

General description of the file including:

**Dimensions of the dataset:**
 
 ```r 
dim(proyect2)

```
 


Summary of the data: 
 
**Activities:**   
 
 ```{r echo=F, results = 'asis'}

panderOptions("digits", 2)

pander(summary(proyect2$activities))

```



##### Values of activities by features:

 ```{r echo=F, results = 'asis'}

panderOptions("digits", 2)

pander(summary(proyect2$value))

```
 
 
 - Variables present in the dataset: 
 
 
 ```{r echo=F, eval=F}
colnames(proyect2)
```



###Variable 1: IDs

This variable indicate the id of the patients of the study.

Some information on the variable including:    
 - Class of the variable: `r class(proyect2$id)`.  
 - Unique values/levels of the variable `r unique(proyect2$id)`.   
 - Unit of measurement: none.   
 - In case names follow some schema, describe how entries were constructed (for example time-body-gyroscope-z has 4 levels of descriptors. Describe these 4 levels). 



####Notes on variable 1:
There are 30 patients.


###Variable 2: activities

This variable indicate the differnt types of activities recorded.

Some information on the variable including:    
 - Class of the variable: `r class(proyect2$activities)`.  
 - Unique values/levels of the variable `r unique(proyect2$activities)` 
 - Unit of measurement: none.   



####Notes on variable 2:
There are 5 different activities recorded.


###Variable 3: features

This variable indicate the different features recorded.

Some information on the variable including:    
 - Class of the variable: `r class(proyect2$features)`.  
 - Unique values/levels of the variable: this is the length of the unique values of the features in the data set: `r length(unique(proyect2$features))`.   
 - Unit of measurement: time and frequency.   
 - In case names follow some schema, describe how entries were constructed (for example time-body-gyroscope-z has 4 levels of descriptors. Describe these 4 levels). 



####Notes on variable 3:
There are `r length(unique(proyect2$features))` factors in the data set.



###Variable 4: value of features

This variable indicates the value of each features in the data set.

Some information on the variable including:    
 - Class of the variable: `r class(proyect2$value)`.  
 - Unique values/levels of the variable `r length(unique(proyect2$value))`.   
 - Unit of measurement: time and frequency.   




####Notes on variable 4:

There are `r length(unique(proyect2$value))` different values in this column.








##Sources

All provided by the course.

##Annex

All the code can be seen in the RScript.